The Talking Heads Experiment

### 2015-04-10 chap12/chap12.tex
cumulative update, mainly for Ch. 7-12

{\bfshape Fitness landscapes of language systems}\\

@@

Here is another example of a fitness landscape from the domain of space, developed by 
Michael Spranger.\footnote{These results are documented at length in: \cite{Spranger:2014}.}
It focuses on understanding the role of grammar for optimising spatial language. \is{Spatial Language Game}
The experiment involves QRIO humanoids
(see \figref{fig:spatial}) and uses a 
sophisticated vision system developed by Spranger, together with IRL and Fluid 
Construction Grammar (in the previous chapter). A variety of spatial language strategies 
were operationalised, in particular:

\begin{figure}
\centerline{\includegraphics[width=0.8\linewidth]{chap11/figs/spatial.pdf}}
\caption{ Set-up for the spatial language game experiments. Two robots are looking at a scene 
with a few blocks, a bigger box with tags indicating front, back or left and right side, and a global landmark in the form 
of a large tag pasted on the wall. {\itshape Top left and right:} World as seen through the camera of the robot. 
{\itshape Bottom left and right:} The internal world model as perceived by the left and right robot.}
\label{fig:spatial}
\end{figure}
\begin{enumerate}
\item The use of {\itshape projective relations} such as front/back, above/below, left/right. This requires imposing a coordinate
system on the world model and name areas within this coordinate system. A common coordinate system is to take 
an object, such as the human body of the speaker as reference point and project vectors emanating from the body. 
\item The use of {\itshape proximal relations} such as far/near. This requires that the distances of objects with respect to the 
reference point are computed and categorised. 
\item The use of {\itshape absolute relations}, such as North/West/East/South, related to global directions. 
\item The use of an {\itshape allocentric landmark} (as opposed to an egocentric one)
from which the situation is conceptualised. For example, in the 
phrase `the block in front of the box', the box is acting as a landmark and the projective spatial relations are 
computed as emanating from that object. Another example is `the block on your left' where the landmark is the 
hearer. 
\end{enumerate}

@@

\begin{figure}
\centerline{\includegraphics[width=0.5\linewidth]{chap12/figs/why-grammar-german.pdf}}
\caption{ Investigation how the two language systems (lexical or grammatical) are performing with respect to the 
communicative success dimension and for each of the three ecological conditions. A purely lexical system in itself has 
already considerable success, but grammar becomes more and more relevant when the situation becomes more complex.
The bottom images show examples for the three ecological conditions being examined.}
\label{fig:why-grammar-german}
\end{figure}

@@

\begin{figure}
\centerline{\includegraphics[width=0.5\linewidth]{chap12/figs/why-grammar-german-avg-interpretations.pdf}
\includegraphics[width=0.5\linewidth]{chap12/figs/why-grammar-conceptualisation-interpretation-equality.pdf}}
\caption{ Left: performance for semantic ambiguity for the same three ecological conditions. We see that without grammar
the world model needs to be consulted heavily to weed out potential interpretations. And Right: Performance for interpretation sharing.
Only with grammar are agents able to share the same semantic structure, ensuring a higher degree of communicative success.}
\label{fig:ambiguity}
\end{figure}

{\bfshape The fitness landscape of language strategies}\\

@@

\begin{figure}[htb!]
\centerline{
\includegraphics[width=0.4\textwidth]{chap12/figs/brightness-graph.pdf}
\includegraphics[height=0.3\textwidth]{chap12/figs/bw-1000.pdf} $~~~$
\includegraphics[height=0.3\textwidth]{chap12/figs/bw-4000.pdf}}
\label{fig:brightness-dynamics}
\caption{{The {\itshape Brightness Colour Strategy} also allows a population of agents (in this case 10)
to evolve an adequate colour lexicon (left). The evolution of a typical lexicon in a smaller population (5 agents), is shown 
after 400 (middle) and 1600 (right) games per agent. Each row represents
the lexicon of one agent.}}
\end{figure}

@@

\begin{figure}[thb!]
\centerline{\includegraphics[width=.75\textwidth]{chap12/figs/strategies-one-winner.pdf}}
\caption{{Communicative success stays constant. The Brightness Colour Strategy becomes 
dominant very early due to random fluctuations. It stays dominant but barely. 
The strategy coherence (i.e. whether agents agree on the 
default strategy) increases gradually. The Full Colour Strategy remains still in use for cases where
it is needed. \label{fig:strategies-one}}}
\end{figure}

@@

\begin{figure}[thb!]
\centerline{\includegraphics[width=.75\textwidth]{chap12/figs/strategies-competition.pdf}}
\caption{{Initially the brightness strategy is dominant, but then the full-colour 
strategy gradually takes over and becomes dominant around 400 games/agents. Strategy coherence within the population
progressively increases.}
\label{fig:strategies-dynamics}}
\end{figure}

@@

{\bfshape A Meta-strategy for Generating new Conceptualisation Strategies}\\

@@

\begin{figure}
\centerline{\includegraphics[width=1.0\linewidth]{chap11/figs/near-the-box2.pdf}}
\caption{ Example of a conceptualisation for the phrase `near to the box'. {\itshape Top:} 
Leftmost: The situation as perceived by the robot. Middle: situation after a geometric transform where the box is 
made the central perspective. Right: Area around the box which reflects (in terms of darkness of the color) how 
near objects are to the box. {\itshape Bottom:} IRL constraint network for this conceptualisation.
\label{fig:spatial-strategies}}
\end{figure}

We will use the domain of spatial language again as a source of examples, based on experiments carried out by 
Michael Spranger. It uses the QRIO humanoid robots 
and scenes like the one shown in \figref{fig:spatial}. An example of a constraint network from this domain 
is shown in \figref{fig:spatial-strategies}. It is the network for the phrase `near the box'. `Near' is a proximal 
relation which categorises objects in terms of how near they are to a landmark (in this case the box). In a list form the 
network looks like this: 
\begin{verbatim}
1. (get-context ?ctx-1) 
2. (construct-region-proximal 
      ?region ?ctx-1 ?landmark ?cat) 
3. (bind proximal-category ?cat NEAR) 
4. (apply-selection ?landmark ?landmarks ?selector-2)
5. (apply-class ?landmarks ?ctx ?landmark-class) 
6. (bind selector ?selector-2 UNIQUE) 
7. (get-context ?ctx) 
8. (bind object-class ?landmark-class BOX)
\end{verbatim}
The order of these operations does not matter as values propagate and constraints become active until no further progress
can be made. Operations in 4, 5, 6, 7 and 8 provide the referent of `the box'. The following contributions 
can be made by each of the cognitive operations: The predicate BOX gets bound to ?landmark-class in operation 8. Object-class is 
the type of ?landmark-class. The context, i.e. all objects in the present situation, gets bound to ?ctx in operation 7. 
The selector UNIQUE gets bound to ?selector-2 in operation 6. 
Then the operation apply-class is applied to all the objects in ?ctx (the context) to find the possible landmarks which 
get bound to ?landmarks using the ?landmark-class (BOX). This set is assumed to be a singleton and so 
in operation 4 the unique element in the set of ?landmarks can be selected 
and bound to ?landmark.

@@

\begin{figure}
\begin{centering}
\includegraphics[width=1\columnwidth]{chap12/figs/conceptualisation.pdf}
\end{centering}
\caption{ The meaning of an utterance is found through a search process, starting from an empty network, progressively more 
constraints are added until a solution network is found that, when executed, yields the topic chosen by the speaker.}
\label{fig:tree}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\columnwidth]{chap12/figs/chunk-alignment-category-invention-success.pdf}
\includegraphics[width=0.8\columnwidth]{chap12/figs/chunk-alignment-category-invention-alignment.pdf}
\caption{ This experiment combines strategy invention, alignment, and category development by each strategy. The top figure
shows language games on the x-axis, and communicative success, average 
number of categories, and interpretation similarity on the y-axis. The bottom figure shows the similarity between 
strategies and the average number of strategies in the population. One strategy becomes dominant.}
\label{fig:all}
\end{center}
\end{figure}

@@

{\bfshape Meta-strategies for generating new lexico-grammatical strategies}\\

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.45\columnwidth]{chap12/figs/attribute-scores-4.pdf}
\includegraphics[width=0.45\columnwidth]{chap12/figs/dog-meaning.pdf}
\end{center}
\caption{ Two examples of the meaning of words in terms of a set of component features whose weight can change 
over time.}
\label{fig:weightchange}
\end{figure}

@@

\begin{figure}
\centerline{i) \includegraphics[scale=.5]{chap12/figs/erosion-performance-summary.pdf}} % fig14-i-erosion-performance
\centerline{ii) \includegraphics[scale=.4]{chap12/figs/uinbui.pdf}} % fig14-ii-uinbui
  \caption{ {Performance of the phonological reduction strategy.} i) Average performance values
for 50 game series for a total of 50,000 games (average of 10,000 per agent). We see that the length of the markers progressively 
diminishes, thus reducing articulatory effort, 
but variation $V_{g}$ does not increase, implying that the system is stable. The marker inventory size remains constant as well.
ii) Trace of the changes to a marker in a single
experiment. The marker {\itshape -uinbui} erodes progressively to {\itshape -u}. A truncated variant is typically present for 
a while in the population until a phase transition happens and it becomes dominant.}
  \label{f:erosion}
\end{figure}

>>>>

\subsection{Fitness landscapes of language systems}
% {\bfshape Fitness landscapes of language systems}\\

@@

Here is another example of a fitness landscape from the domain of space, developed by 
Michael Spranger.\footnote{These results are documented at length in: \cite{Spranger:2014}.}
It focuses on understanding the role of grammar for optimising spatial language. \is{Spatial Language Game}
The experiment involves \textsc{qrio} humanoids
(see \figref{fig:spatial}) and uses a 
sophisticated vision system developed by Spranger, together with IRL and Fluid 
Construction Grammar (in the previous chapter). A variety of spatial language strategies 
were operationalised, in particular:

\begin{figure}
\centerline{\includegraphics[width=0.8\linewidth]{chap11/figs/spatial.pdf}}
\caption{Set-up for the spatial language game experiments. Two robots are looking at a scene 
with a few blocks, a bigger box with tags indicating front, back or left and right side, and a global landmark in the form 
of a large tag pasted on the wall. {\itshape Top left and right:} World as seen through the camera of the robot. 
{\itshape Bottom left and right:} The internal world model as perceived by the left and right robot.}
\label{fig:spatial}
\end{figure}
\begin{enumerate}
\item The use of {\itshape projective relations} such as front/back, above/below, left/right. This requires imposing a coordinate
system on the world model and name areas within this coordinate system. A common coordinate system is to take 
an object, such as the human body of the speaker as reference point and project vectors emanating from the body. 
\item The use of {\itshape proximal relations} such as far/near. This requires that the distances of objects with respect to the 
reference point are computed and categorised. 
\item The use of {\itshape absolute relations}, such as North/West/East/South, related to global directions. 
\item The use of an {\itshape allocentric landmark} (as opposed to an egocentric one)
from which the situation is conceptualised. For example, in the 
phrase `the block in front of the box', the box is acting as a landmark and the projective spatial relations are 
computed as emanating from that object. Another example is `the block on your left' where the landmark is the 
hearer. 
\end{enumerate}

@@

\begin{figure}
\centerline{\includegraphics[width=0.5\linewidth]{chap12/figs/why-grammar-german.pdf}}
\caption{Investigation how the two language systems (lexical or grammatical) are performing with respect to the 
communicative success dimension and for each of the three ecological conditions. A purely lexical system in itself has 
already considerable success, but grammar becomes more and more relevant when the situation becomes more complex.
The bottom images show examples for the three ecological conditions being examined.}
\label{fig:why-grammar-german}
\end{figure}

@@

\begin{figure}
\centerline{\includegraphics[width=0.5\linewidth]{chap12/figs/why-grammar-german-avg-interpretations.pdf}
\includegraphics[width=0.5\linewidth]{chap12/figs/why-grammar-conceptualisation-interpretation-equality.pdf}}
\caption{Left: performance for semantic ambiguity for the same three ecological conditions. We see that without grammar
the world model needs to be consulted heavily to weed out potential interpretations. And Right: Performance for interpretation sharing.
Only with grammar are agents able to share the same semantic structure, ensuring a higher degree of communicative success.}
\label{fig:ambiguity}
\end{figure}

\subsection{The fitness landscape of language strategies}
% {\bfshape The fitness landscape of language strategies}\\

@@

\begin{figure}[htb!]
\centerline{
\includegraphics[width=0.4\textwidth]{chap12/figs/brightness-graph.pdf}
\includegraphics[height=0.3\textwidth]{chap12/figs/bw-1000.pdf} $~~~$
\includegraphics[height=0.3\textwidth]{chap12/figs/bw-4000.pdf}}
\label{fig:brightness-dynamics}
\caption{{The {\itshape Brightness Colour Strategy} also allows a population of agents (in this case 10)
to evolve an adequate colour lexicon (left). The evolution of a typical lexicon in a smaller population (5 agents), is shown 
after 400 (middle) and 1600 (right) games per agent. Each row represents
the lexicon of one agent.}}\todo[inline]{barely readable at 100\%. Do we need to change s.th. here?}
\end{figure}

@@

\begin{figure}[thb!]
\centerline{\includegraphics[width=\textwidth]{chap12/figs/strategies-one-winner.pdf}}
\caption{{Communicative success stays constant. The Brightness Colour Strategy becomes 
dominant very early due to random fluctuations. It stays dominant but barely. 
The strategy coherence (i.e. whether agents agree on the 
default strategy) increases gradually. The Full Colour Strategy remains still in use for cases where
it is needed. \label{fig:strategies-one}}}
\end{figure}

@@

\begin{figure}[thb!]
\centerline{\includegraphics[width=\textwidth]{chap12/figs/strategies-competition.pdf}}
\caption{{Initially the brightness strategy is dominant, but then the full-colour 
strategy gradually takes over and becomes dominant around 400 games/agents. Strategy coherence within the population
progressively increases.}
\label{fig:strategies-dynamics}}
\end{figure}

@@

\subsection{A meta-strategy for generating new conceptualisation strategies}
% {\bfshape A Meta-strategy for Generating new Conceptualisation Strategies}\\

@@

\begin{figure}
\centerline{\includegraphics[width=1.0\linewidth]{chap11/figs/near-the-box2.pdf}}
\caption{Example of a conceptualisation for the phrase `near to the box'. {\itshape Top:} 
Leftmost: The situation as perceived by the robot. Middle: situation after a geometric transform where the box is 
made the central perspective. Right: Area around the box which reflects (in terms of darkness of the color) how 
near objects are to the box. {\itshape Bottom:} IRL constraint network for this conceptualisation.
\label{fig:spatial-strategies}}
\end{figure}

We will use the domain of spatial language again as a source of examples, based on experiments carried out by 
Michael Spranger. It uses the \textsc{qrio} humanoid robots 
and scenes like the one shown in \figref{fig:spatial}. An example of a constraint network from this domain 
is shown in \figref{fig:spatial-strategies}. It is the network for the phrase `near the box'. `Near' is a proximal 
relation which categorises objects in terms of how near they are to a landmark (in this case the box). In a list form the 
network looks like this: 
\begin{verbatim}
1. (get-context ?ctx-1) 
2. (construct-region-proximal 
      ?region ?ctx-1 ?landmark ?cat) 
3. (bind proximal-category ?cat NEAR) 
4. (apply-selection ?landmark ?landmarks ?selector-2)
5. (apply-class ?landmarks ?ctx ?landmark-class) 
6. (bind selector ?selector-2 UNIQUE) 
7. (get-context ?ctx) 
8. (bind object-class ?landmark-class BOX)
\end{verbatim}
The order of these operations does not matter as values propagate and constraints become active until no further progress
can be made. Operations in 4, 5, 6, 7 and 8 provide the referent of `the box'. The following contributions 
can be made by each of the cognitive operations: The predicate BOX gets bound to ?landmark-class in operation 8. Object-class is 
the type of ?landmark-class. The context, i.e. all objects in the present situation, gets bound to ?ctx in operation 7. 
The selector UNIQUE gets bound to ?selector-2 in operation 6. 
Then the operation apply-class is applied to all the objects in ?ctx (the context) to find the possible landmarks which 
get bound to ?landmarks using the ?landmark-class (BOX). This set is assumed to be a singleton and so 
in operation 4 the unique element in the set of ?landmarks can be selected 
and bound to ?landmark.

@@

\begin{figure}
\begin{centering}
\includegraphics[width=1\columnwidth]{chap12/figs/conceptualization.pdf}
\end{centering}
\caption{The meaning of an utterance is found through a search process, starting from an empty network, progressively more 
constraints are added until a solution network is found that, when executed, yields the topic chosen by the speaker.}
\label{fig:tree}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{chap12/figs/chunk-alignment-category-invention-success.pdf}
\includegraphics[width=\textwidth]{chap12/figs/chunk-alignment-category-invention-alignment.pdf}
\caption{This experiment combines strategy invention, alignment, and category development by each strategy. The top figure
shows language games on the x-axis, and communicative success, average 
number of categories, and interpretation similarity on the y-axis. The bottom figure shows the similarity between 
strategies and the average number of strategies in the population. One strategy becomes dominant.}
\label{fig:all}
\end{center}
\end{figure}

@@

\subsection{Meta-strategies for generating new lexico-grammatical strategies}
% {\bfshape Meta-strategies for generating new lexico-grammatical strategies}\\

@@

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{chap12/figs/attribute-scores-4.pdf}
\includegraphics[width=\textwidth]{chap12/figs/dog-meaning.pdf}
\end{center}
\caption{Two examples of the meaning of words in terms of a set of component features whose weight can change over time.}
\label{fig:weightchange}
\end{figure}

@@

\begin{figure}
\centerline{i) \includegraphics[scale=.5]{chap12/figs/erosion-performance-summary.pdf}} % fig14-i-erosion-performance
\centerline{ii) \includegraphics[scale=.4]{chap12/figs/uinbui.pdf}} % fig14-ii-uinbui
  \caption{{Performance of the phonological reduction strategy.} i) Average performance values
for 50 game series for a total of 50,000 games (average of 10,000 per agent). We see that the length of the markers progressively 
diminishes, thus reducing articulatory effort, 
but variation $V_{g}$ does not increase, implying that the system is stable. The marker inventory size remains constant as well.
ii) Trace of the changes to a marker in a single
experiment. The marker {\itshape -uinbui} erodes progressively to {\itshape -u}. A truncated variant is typically present for 
a while in the population until a phase transition happens and it becomes dominant.}
  \label{f:erosion}
\end{figure}

### 2015-04-10 chap5/chap5.tex
cumulative update, mainly for Ch. 7-12

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.60\textwidth]{chap5/figs/incr-decr.pdf}}
\caption{\label{incr-decr} 
Score adjustments after a successful game. Scores of used 
associations go up and their competitors go down.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=\textwidth]{chap5/figs/success.pdf}}
\caption{\label{success} 
The graph displays on the y-axis the average success every 10 games
in a population of five agents lexicalising five
different meanings. The x-axis 
shows the number of games. Average success
rapidly climbs until it reaches total success after 
about 180 games.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=\textwidth]{chap5/figs/larger.pdf}}
\caption{\label{larger} 
The graph displays the evolution of communicative success
for larger and larger populations. The
number of games on the x-axis
is divided by the number of agents. }
\end{figure}
We see for example that for 20 agents and 20 meanings 
success climbs to total success after about 10,000 games.
This is still surprisingly low particularly as
success is
are already above 95 \% after about 5000 games. Games can be played
in parallel by different agents because the system 
is entirely distributed. If we divide
the number of agents by the number of games, we see that about 250
games are needed by the agents to get 95 \% success, which means
that every meaning needs to appear about 10 times for each agent. 
Interestingly enough, the larger the population of agents, 
the more the success curve approximates an 
S-shape, which has 
been observed empirically in the spreading of new
linguistic conventions. The same curve shape is familiar to
biologists studying models of competitive growth, suggesting 
a strong relationships between ecological dynamics and language\is{S-shape curve}
spreading.\footnote{The S-shaped curve is discussed in \cite{McMahon:1994} p. 52. For examples
of biological models with similar properties, see \cite{May:1976}.}

>>>>

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.60\textwidth]{chap5/figs/incr-decr.pdf}}
\caption{\label{incr-decr}Score adjustments after a successful game. Scores of used 
associations go up and their competitors go down.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=\textwidth]{chap5/figs/success.pdf}}
\caption{\label{success}The graph displays on the y-axis the average success every 10 games
in a population of five agents lexicalising five
different meanings. The x-axis 
shows the number of games. Average success
rapidly climbs until it reaches total success after 
about 180 games.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=\textwidth]{chap5/figs/larger.pdf}}
\caption{\label{larger}The graph displays the evolution of communicative success
for larger and larger populations. The
number of games on the x-axis
is divided by the number of agents. }
\end{figure}
We see for example that for 20 agents and 20 meanings 
success climbs to total success after about 10,000 games.
This is still surprisingly low particularly as
success is
are already above 95 \% after about 5000 games. Games can be played
in parallel by different agents because the system 
is entirely distributed. If we divide
the number of agents by the number of games, we see that about 250
games are needed by the agents to get 95 \% success, which means
that every meaning needs to appear about 10 times for each agent. 
Interestingly enough, the larger the population of agents, 
the more the success curve approximates an 
S-shape, which has 
been observed empirically in the spreading of new
linguistic conventions. The same curve shape is familiar to
biologists studying models of competitive growth, suggesting 
a strong relationships between ecological dynamics and language\is{S-shape curve}
spreading.\footnote{The S-shaped curve is discussed in \cite{McMahon:1994} p. 52. For examples
of biological models with similar properties, see \cite{May:1976}.}

### 2015-04-10 chap8/chap8.tex
cumulative update, mainly for Ch. 7-12

\begin{figure}[htbp]
  \centerline{\includegraphics[width=0.7\textwidth]{chap8/figures/hu-obrist+ls.pdf}}
\caption{ \label{fig:faces} 
Hans-Ulrich Obrist and Luc Steels during the opening of the Laboratorium Exhibition in Antwerp, summer 1999. }
\end{figure}
In addition to the installation at the exhibition in Antwerp itself, there were two additional 
external sites for the Talking Heads experiment operating within the same time frame: The Paris site at the Sony Computer 
Science Laboratory in Paris and the Brussels site at the Free University (Vrije Universiteit) of Brussels.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=0.9\textwidth]{chap8/figures/layout.pdf}}
\caption{\label{fig:layout} 
Layout of the 'Laboratory for Cognitive Robotics and Teleporation" at the Laboratorium exhibition in Antwerp.}
\end{figure}

\begin{figure}[htbp]
 \centerline{\includegraphics[width=.35\textwidth]{chap8/figures/zoom-in-head.pdf}}
\caption{ \label{fig:laboratorium} 
A single Talking Head at the Paris site. In the background we see (bottom) the display with an
image that the agent picked up, (on top of that) a screen displaying the processed image and the discrimination trees, 
and to the right the loudspeaker broadcasting the speech output of the agent.}
\end{figure}

@@

\begin{figure}[htbp]
 \centerline{\includegraphics[width=.40\textwidth]{chap8/figures/image-1.jpg}
\includegraphics[width=.40\textwidth]{chap8/figures/image-2.pdf}}
\caption{\label{fig:images} 
Examples of images picked up by the camera's. The chosen topic was signalled by drawing a bounding box around it. 
During experimentation, it was found that configurations such as in the right image lead to more stable performance.} 
\end{figure}

@@

\begin{figure}[htbp]
 \centerline{\includegraphics[width=.80\textwidth]{chap8/figures/xagadu-speaker.pdf}}
\centerline{\includegraphics[width=.80\textwidth]{chap8/figures/xagadu-hearer.pdf}}
\caption{\label{fig:interaction} 
Top: Source image, conceptualisation, and word choice of the speaker. Bottom: Source image, parsing and interpretation 
of the hearer.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.85\textwidth]{chap8/figures/th-website.pdf}}
\caption{\label{fig:thwebsite} 
The Talking Heads website main interface. Users could create and manage their own agents and inspect what was happening 
at each site. Here 12 agents located at the Brussels site are listed.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.85\textwidth]{chap8/figures/zelebot.pdf}}
\caption{\label{fig:thwebsite-agent} 
Example of an interaction that took part in Antwerp. This game was a success. The agents recognised the same
set of objects and the meaning of `kazozo' was effective in finding the same topic. 
}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.80\textwidth]{chap8/figures/allsites.pdf}}
\caption{\label{fig:allsites} 
The y-axis shows the average communicative success of agents at each of the three different sites for a total 
of 190.000 games shown on the y-axis. 
}
\end{figure}

@@

\begin{figure}[htbp]
 \centerline{\includegraphics[width=.80\textwidth]{chap8/figures/words-used.pdf}}
\caption{\label{fig:words-used} 
Distribution of word use. There are only a few words that are dominant. Many words are 
short-lived, either because the circumstances 
in which they fit are rare or because they do not get settled in the population. Moreover sometimes users create agents 
which they do not keep rescheduling for new games. 
}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.80\textwidth,height= 6cm]{chap8/figures/wogglesplat.pdf}}
\caption{\label{fig:woggle} 
Meaning-Form diagram: Different words for expressing the meaning 'left', i.e. 
horizontal position is less than 0.25 (scaled). 
New words come up all the time but there is a clear winner-take-all effect with "wogglesplat" winning. 
}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.80\textwidth,height=6cm]{chap8/figures/droite.pdf}}
\caption{\label{fig:droite} 
Form-Meaning diagram: Different meanings associated with the word `droite'. The dominant meaning is 
in line with the human use of the word, namely horizontal position (HPOS) to the right of the image.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.70\textwidth, height=6cm]{chap8/figures/bozopite.pdf}}
\caption{\label{fig:bozopite} 
Form-Meaning diagram: Different meanings associated with the word `bozopite'. 
}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.65\textwidth]{chap8/figures/down.pdf}}
\caption{\label{fig:down} 
Form-Meaning diagram: Different meanings associated with the word "down". There are in particular two meanings. 
one is a region on the red-channel and another one a region on the vertical position channel. 
}
\end{figure}

>>>>

\begin{figure}[htbp]
  \centerline{\includegraphics[width=0.7\textwidth]{chap8/figures/hu-obrist+ls.pdf}}
\caption{\label{fig:faces}Hans-Ulrich Obrist and Luc Steels during the opening of the Laboratorium Exhibition in Antwerp, summer 1999. }
\end{figure}
In addition to the installation at the exhibition in Antwerp itself, there were two additional 
external sites for the Talking Heads experiment operating within the same time frame: The Paris site at the Sony Computer 
Science Laboratory in Paris and the Brussels site at the Free University (Vrije Universiteit) of Brussels.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=0.9\textwidth]{chap8/figures/layout.pdf}}
\caption{\label{fig:layout}Layout of the 'Laboratory for Cognitive Robotics and Teleporation" at the Laboratorium exhibition in Antwerp.}
\end{figure}

\begin{figure}[htbp]
 \centerline{\includegraphics[width=.35\textwidth]{chap8/figures/zoom-in-head.pdf}}
\caption{\label{fig:laboratorium}A single Talking Head at the Paris site. In the background we see (bottom) the display with an
image that the agent picked up, (on top of that) a screen displaying the processed image and the discrimination trees, 
and to the right the loudspeaker broadcasting the speech output of the agent.}
\end{figure}

@@

\begin{figure}[htbp]
 \centerline{\includegraphics[width=.40\textwidth]{chap8/figures/image-1.jpg}
\includegraphics[width=.40\textwidth]{chap8/figures/image-2.pdf}}
\caption{\label{fig:images}Examples of images picked up by the camera's. The chosen topic was signalled by drawing a bounding box around it. 
During experimentation, it was found that configurations such as in the right image lead to more stable performance.} 
\end{figure}

@@

\begin{figure}[htbp]
 \centerline{\includegraphics[width=.80\textwidth]{chap8/figures/xagadu-speaker.pdf}}
\centerline{\includegraphics[width=.80\textwidth]{chap8/figures/xagadu-hearer.pdf}}
\caption{\label{fig:interaction}Top: Source image, conceptualisation, and word choice of the speaker. Bottom: Source image, parsing and interpretation 
of the hearer.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.85\textwidth]{chap8/figures/th-website.pdf}}
\caption{\label{fig:thwebsite}The Talking Heads website main interface. Users could create and manage their own agents and inspect what was happening 
at each site. Here 12 agents located at the Brussels site are listed.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.85\textwidth]{chap8/figures/zelebot.pdf}}
\caption{\label{fig:thwebsite-agent}Example of an interaction that took part in Antwerp. This game was a success. The agents recognised the same
set of objects and the meaning of `kazozo' was effective in finding the same topic. 
}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=\textwidth]{chap8/figures/allsites.pdf}}
\caption{\label{fig:allsites}The y-axis shows the average communicative success of agents at each of the three different sites for a total 
of 190.000 games shown on the y-axis. 
}
\end{figure}

@@

\begin{figure}[htbp]
 \centerline{\includegraphics[width=\textwidth]{chap8/figures/words-used.pdf}}
\caption{\label{fig:words-used}Distribution of word use. There are only a few words that are dominant. Many words are 
short-lived, either because the circumstances 
in which they fit are rare or because they do not get settled in the population. Moreover sometimes users create agents 
which they do not keep rescheduling for new games. 
}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=\textwidth]{chap8/figures/wogglesplat.pdf}}
\caption{\label{fig:woggle}Meaning-Form diagram: Different words for expressing the meaning 'left', i.e. 
horizontal position is less than 0.25 (scaled). 
New words come up all the time but there is a clear winner-take-all effect with "wogglesplat" winning. 
}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=\textwidth]{chap8/figures/droite.pdf}}
\caption{\label{fig:droite}Form-Meaning diagram: Different meanings associated with the word `droite'. The dominant meaning is 
in line with the human use of the word, namely horizontal position (HPOS) to the right of the image.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=\textwidth]{chap8/figures/bozopite.pdf}}
\caption{\label{fig:bozopite}Form-Meaning diagram: Different meanings associated with the word `bozopite'. 
}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.65\textwidth]{chap8/figures/down.pdf}}
\caption{\label{fig:down}Form-Meaning diagram: Different meanings associated with the word "down". There are in particular two meanings. 
one is a region on the red-channel and another one a region on the vertical position channel. 
}
\end{figure}
